The \dealii Library, Version 9.6

### 2024-09-17 9.6/paper.tex
Write text

The major changes of this release are:
%
\begin{itemize}
  \item Substantial improvements to the multigrid two-level transfer
    operators, including matrix-free based versions and versions for
    p-refinement (Section~\ref{sec:mf}).
  \item Different additions to the non-matching infrastructure (see
    Section~\ref{sec:nonmatching}) In particular, the new class
    FERemoteEvaluation provides support for evaluating
    finite element shape functions and solutions on parts of the
    domain stored by other MPI processes.
  \item Much work has gone into writing wrappers for Trilinos' Tpetra
    stack of linear algebra classes. Tpetra is Trilinos' Kokkos-based
    replacement for the now-deprecated Epetra stack. See
    Section~\ref{sec:tpetra} for more on this.
  \item Tool classes \texttt{TaskResult} and \texttt{Lazy} that provide ways to compute
    values on a separate thread or only when first accessed (Section~\ref{sec:tools}).
  \item There are six new tutorial programs, on checkpointing simulations
    (step-83), integrating time-dependent solvers with external time
    stepping libraries (step-86, using PETSc's TS library), advanced
    point evaluation techniques (step-87), non-matching grids
    (step-89), and trace-based methods for PDEs on embedded surfaces
    (step-90). See Section~\ref{subsec:steps} for more details.
\end{itemize}
%

@@

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item deal.II now requires and uses of C++17.
  \item We have continued to make progress in supporting simplex and
    mixed meshes -- mesh types that \dealii{} has traditionally not
    supported at all. Specifically, the current release uses
    better strategies for refinement of tetrahedra that result in
    better-shaped child cells. It also contains
    support for cubic finite elements on simplices.
  \item The \texttt{FE\_NedelecSZ} class that contains our
    implementation of the N\'ed\'elec element using the orientation
    scheme of \cite{Zag06} now supports the computation of hanging
    node constraints for locally refined meshes.
    \todo[inline]{Sebastian: Is there anything more to say here, maybe some reference? If
      you want, we can also list this point in the list above, and add
      a whole section 2.x for it.}
  \item changes of interface of AffineConstraints
    \todo[inline]{Wolfgang to write}
\item clean up: orientation
  \todo[inline]{DavidW: Please describe some details here.}
\item Various performance improvements throughout the library.
  \todo[inline]{Everyone: This sentence would be more useful if we can
    provide a few examples.}
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.

@@

\todo[inline]{Martin/Peter: The list in the introduction also mentions the two-level
  operators. Is this the right place to also discuss these?}

\begin{itemize}
\item VectorizedArray for ARM Neon
\item Improvement for evaluation of values and gradients for
  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements
\item Improved internal data structures of the tensor-product evaluators as
  well as the evaluators for simplex elements, which speed up the evaluation
  in several scenarios, especially multi-component systems.
\end{itemize}

@@

\todo[inline]{All: If you contributed to this functionality, please
  help flesh out this section.}

\begin{itemize}
\item non-nested multigrid: new features and performance improvement
\item FERemoteEvaluation
\item FECouplingValues
\end{itemize}

@@

>>>>

The major changes of this release are:
%
\begin{itemize}
  \item Substantial improvements to the multigrid two-level transfer
    operators, including matrix-free based versions and versions for
    p-refinement (Section~\ref{sec:mf}).
  \item Different additions to the non-matching infrastructure (see
    Section~\ref{sec:nonmatching}). In particular, the new class
    FERemoteEvaluation provides support for evaluating
    finite element shape functions and solutions on parts of the
    domain stored by other MPI processes.
  \item Much work has gone into writing wrappers for Trilinos' Tpetra
    stack of linear algebra classes. Tpetra is Trilinos' Kokkos-based
    replacement for the now-deprecated Epetra stack. See
    Section~\ref{sec:tpetra} for more on this.
  \item Tool classes \texttt{TaskResult} and \texttt{Lazy} that provide ways to compute
    values on a separate thread or only when first accessed (Section~\ref{sec:tools}).
  \item There are six new tutorial programs, on checkpointing simulations
    (step-83), integrating time-dependent solvers with external time
    stepping libraries (step-86, using PETSc's TS library), advanced
    point evaluation techniques (step-87), non-matching grids
    (step-89), and trace-based methods for PDEs on embedded surfaces
    (step-90). See Section~\ref{subsec:steps} for more details.
\end{itemize}
%

@@

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item deal.II now requires and uses of C++17.
  \item We have continued to make progress in supporting simplex and
    mixed meshes -- mesh types that \dealii{} has traditionally not
    supported at all. Specifically, the current release uses
    better strategies for refinement of tetrahedra that result in
    better-shaped child cells. It also contains
    support for cubic finite elements on simplices.
  \item The \texttt{FE\_NedelecSZ} class that contains our
    implementation of the N\'ed\'elec element using the orientation
    scheme of \cite{Zag06} now supports the computation of hanging
    node constraints for locally refined meshes.
    \todo[inline]{Sebastian: Is there anything more to say here, maybe some reference? If
      you want, we can also list this point in the list above, and add
      a whole section 2.x for it.}
  \item changes of interface of AffineConstraints
    \todo[inline]{Wolfgang to write}
\item We have clean up the orientation definition of faces in (2D and) 3D. Three
    Booleans used to be used to define it for hexedras: \textit{orientation},
    \textit{flip}, and \textit{rotation}. While introducing simplex and mixed
     mesh support in \dealii, we merged these three into a single variable of type
     char internally. Now, we also expose these to the outside, leading to
     multiple incompatibilities, since the template type of different data
     structures has changed and the number of input/output arguments have decreased.
  \todo[inline]{DavidW: Please describe some details here.}
\item We have performed various performance improvements throughout the library,
     which we detail in the following sections.
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.

@@

%\todo[inline]{Martin/Peter: The list in the introduction also mentions the two-level
%  operators. Is this the right place to also discuss these?}

We made different updates to the matrix-free infrastructure in \dealii. These include:
\begin{itemize}
\item Our own implementation of \texttt{std::simd} (\texttt{VectorizedArray}) now
  also Arm Neon. Arm Neon is an architecture extension for
  the Arm Cortex-A and Arm Cortex-R series of processors, commonly used in Apple
  products. With these instructions 2 doubles or 4 floats can be processed in one
  go. Since matrix-free infrastructure works directly with \texttt{VectorizedArray}
  data structures, it automatically benefits from this extension.

\item We made some improvement for evaluation of values and gradients for
  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements.

\item We furthermore improved the internal data structures of the
  tensor-product evaluators as well as the evaluators for simplex elements,
  which speed up the evaluation in several scenarios, especially multi-component systems.
\end{itemize}

Furthermore, we performed substantial improvements to the global-coarsening
multigrid infrastructure: \texttt{MGTransferMF} (previously:
\texttt{MGTransferGlobalCoarsening}) and \texttt{MGTwoLevelTransfer}. They
now allow to perform local smoothing, which is a key step towards unifying
all transfer operators in \dealii. \texttt{MGTwoLevelTransfer} now also
can be set up with an existing \texttt{MatrixFree} object in the case
of $p$-multigrid, allowing to reduce the setup costs significantly.

@@

%\todo[inline]{All: If you contributed to this functionality, please
%  help flesh out this section.}

We made different advances to the non-matching infrastructure in \dealii. These include:

\begin{itemize}
\item The performance of the non-nested multigrid infrastructure
(\texttt{MGTwoLevelTransferNon\-Nested}) has been improved significantly by avoiding
redundant copy operations. Furthermore, the support for simplex-shaped cells and
multiple-component elements have been added.
\item The new class \texttt{FERemoteEvaluation} has been added. 
This is a class to access data
in a distributed matrix-free loop for non-matching discretizations.
Interfaces are named with \texttt{FEEvaluation} in mind. Key component is the
underlying MPI communication, which is done via \texttt{RemotePointEvaluation}.
Tutorial \texttt{step-89} has been added to present its usage
in the context of an application to acoustic conservation equations~\cite{heinz2023high}.
\item FECouplingValues

\todo[inline]{Luca: Please complete}
\end{itemize}

@@

\todo[inline]{Peter: Anything to list here?}
